---
title: "Grade analysis lab"
author: "Miriam Schäfer"
format: 
  html:
    echo: false 
    warning: false
---

### Question 1
*Load the data set using a local file name (using here). Notice that the file is included in the forked repository.*

```{r}
here::i_am("r-101-grade-analysis.Rproj")
library(ggplot2)
library(dplyr)
library(tidyr) 

grades <- read.csv("grades.csv")
```


### Question 2
*Compute the minimum, maximum, median and mean grade at the Exam and include the results as a table in the rendered document (using, e.g. knitr::kable).*


```{r}
q2 <- grades |> 
  filter(!is.na(Exam)) |>
  summarize(
    Min = min(Exam), 
    Max = max(Exam), 
    Median = median(Exam),
    Mean = mean(Exam)
  ) 

knitr::kable((q2), "simple")  
```



### Question 3 
*Counts the number of students who did not take the final exam (i.e. for whom Exam is NA) and include the results in a sentence in your rendered document.* 

```{r}
q3 <- grades |> count(is.na(Exam)==TRUE)
```
There were `r q3[2,2]` students who did not take the final exam. 


### Question 4
*Represent graphically in an adapted way the distribution of grades at the Exam. Make sure to adapt the code to avoid any error or warning from ggplot2.*

```{r}
q4 <- grades |> 
  filter(!is.na(Exam)) |>
  ggplot(aes(x=Exam)) + 
  geom_bar(fill = "black", color = "white") +
  labs(y = "Number of students", x = "Exam grade", title ="Distribution of grades")

q4
```

### Question 5 
*Compute the number of students in each Group and include the results in your document as a table.

```{r}
q5 <- grades |> 
  group_by(Group) |>
  summarize(N = n()) 

knitr::kable((q5), "simple")  
```




